Alone Together: Compositional Reasoning and Inference for Weak Isolation

### 2017-10-18 supplement.tex
Edits
Minor Changes in RG-rules

%
\textbf{Top-Level Reasoning} \quad \fbox {\(
\rg{I,R}{c}{G,I} \)}\\
%
%
\begin{minipage}{3.5in}
\rulelabel{RG-Txn}
\begin{smathpar}
\begin{array}{c}
\RULE
{
 \stable(R, I) \spc
  \stable(R,\I)\spc
  P(\stl,\stg) \Leftrightarrow \stl=\emptyset \wedge I(\stg)\\
  \R_e(\stl,\stg,\stg') \Leftrightarrow \exists \stg_1. \I_e(\stl, \stg_1, \stg) \wedge R(\stg, \stg') \wedge \I_e(\stl, \stg_1, \stg') \spc
  \R_e \vdash \rg{P}{c}{Q}\\
  \R_c(\stl,\stg,\stg') \Leftrightarrow \exists \stg_1. \I_c(\stl, \stg_1, \stg) \wedge R(\stg, \stg') \wedge \I_c(\stl, \stg_1, \stg')  \spc
  \stable(\R_c,Q)\\
  \forall \stl,\stg.~ Q(\stl,\stg) \Rightarrow 
    G(\stg, \stl \gg \stg)\spc
  \forall \stg,\stg'.~I(\stg) \wedge G(\stg,\stg') \Rightarrow I(\stg')\\
}
{
  \rg{I,R}{\ctxn{i}{\I}{c}}{G \cup ID,I}
}
\end{array}
\end{smathpar}
\end{minipage}
%

@@

\begin{theorem}
RG-Txn is sound.
\end{theorem}
\begin{proof}
\begin{mathpar}
\begin{array}{lc}
  \stable(R, I) & HI\\
  \stable(R,\I) & H\I \\
  P(\stl,\stg) \Leftrightarrow \stl=\emptyset \wedge I(\stg) & HP\\ 
  \R_e(\stl,\stg,\stg') \Leftrightarrow \exists \stg_1. \I_e(\stl, \stg_1, \stg) \wedge R(\stg, \stg') \wedge \I_e(\stl, \stg_1, \stg') & H\R_l \\
  \R_e\vdash \rg{P}{c}{Q} & Hc\\
  \R_c(\stl,\stg,\stg') \Leftrightarrow \exists \stg_1. \I_c(\stl, \stg_1, \stg) \wedge R(\stg, \stg') \wedge \I_c(\stl, \stg_1, \stg') & H\R_c \\
  \stable(\R_c,Q) & HQ \\
  \forall \stl,\stg.~ Q(\stl,\stg) \Rightarrow 
    G(\stg, \stl \gg \stg)\spc  & HQG \\
  \forall \stg,\stg'.~I(\stg) \wedge G(\stg,\stg') \Rightarrow I(\stg') & HG\\
  \forall \stg. G(\stg, \stg) & HID
\end{array}
\end{mathpar}

@@

Now, we will show that every environment step in $\pi$ is in $\R_e$. Assume that $(t_1, \stg_a) \rightarrow_R (t_1, \stg_b)$ is an environment step such that $R(\stg_a, \stg_b)$. Then, we know that $\I_e(t_1.\stl, t_1.\stg, \stg_a)$ and $\I_e(t_1.\stl, t_1.\stg, \stg_b)$. Hence, $t_1.\stg$ provides the existence of $\stg_1$ in the definition of $\R_e$. Thus, $\R_e(t_1.\stl, \stg_a, \stg_b)$. We can use $Hc$ and make the assertion $Q(t^{'}.\stl, \stg_2)$.

@@

Now, we will show that every environment step is in $\R_c$. Consider the step $(t^{'}, \stg_{2k}) \rightarrow_R (t^{'}, \stg_{2(k+1)})$. We have $\I_c(t^{'}.\stl, \stg_2, \stg_{2k})$ and $\I_c(t^{'}.\stl, \stg_2, \stg_{2(k+1)})$. Hence, $\stg_2$ provides the existence of $\stg_1$ in the definition of $\R_c$. Thus, $\R_c(t^{'}.\stl, \stg_{2k}, \stg_{2(k+1)})$.

@@

>>>>

%
\textbf{Top-Level Reasoning} \quad \fbox {\(
\rg{I,R}{c}{G,I} \)}\\
%
%
\begin{minipage}{3.5in}
\rulelabel{RG-Txn}
\begin{smathpar}
\begin{array}{c}
\RULE
{
 \stable(R, I) \spc
  \stable(R,\I)\spc
  P(\stl,\stg) \Leftrightarrow \stl=\emptyset \wedge I(\stg)\\
  \R_e(\stl,\stg,\stg') \Leftrightarrow \exists \stg_1. R(\stg, \stg') \wedge \I_e(\stl, \stg_1, \stg) \wedge \I_e(\stl, \stg_1, \stg') \spc
  \R_e \vdash \rg{P}{c}{Q}\\
  \R_c(\stl,\stg,\stg') \Leftrightarrow \exists \stg_1. R(\stg, \stg') \wedge \I_c(\stl, \stg_1, \stg) \wedge \I_c(\stl, \stg_1, \stg')  \spc
  \stable(\R_c,Q)\\
  \forall \stl,\stg.~ Q(\stl,\stg) \Rightarrow 
    G(\stg, \stl \gg \stg)\spc
  \forall \stg,\stg'.~I(\stg) \wedge G(\stg,\stg') \Rightarrow I(\stg')\\
}
{
  \rg{I,R}{\ctxn{i}{\I}{c}}{G \cup ID,I}
}
\end{array}
\end{smathpar}
\end{minipage}
%

@@

\begin{theorem}
RG-Txn is sound.
\end{theorem}
\begin{proof}
\begin{mathpar}
\begin{array}{lc}
  \stable(R, I) & HI\\
  \stable(R,\I) & H\I \\
  P(\stl,\stg) \Leftrightarrow \stl=\emptyset \wedge I(\stg) & HP\\ 
  \R_e(\stl,\stg,\stg') \Leftrightarrow \exists \stg_1. R(\stg, \stg') \wedge \I_e(\stl, \stg_1, \stg) \wedge  \I_e(\stl, \stg_1, \stg') & H\R_l \\
  \R_e\vdash \rg{P}{c}{Q} & Hc\\
  \R_c(\stl,\stg,\stg') \Leftrightarrow \exists \stg_1. R(\stg, \stg') \wedge \I_c(\stl, \stg_1, \stg) \wedge  \I_c(\stl, \stg_1, \stg') & H\R_c \\
  \stable(\R_c,Q) & HQ \\
  \forall \stl,\stg.~ Q(\stl,\stg) \Rightarrow 
    G(\stg, \stl \gg \stg)\spc  & HQG \\
  \forall \stg,\stg'.~I(\stg) \wedge G(\stg,\stg') \Rightarrow I(\stg') & HG\\
  \forall \stg. G(\stg, \stg) & HID
\end{array}
\end{mathpar}

@@

Now, we will show that every environment step in $\pi$ is in $\R_e$. Assume that $(t_1, \stg_a) \rightarrow_R (t_1, \stg_b)$ is an environment step such that $R(\stg_a, \stg_b)$. Then, we know that $\I_e(t_1.\stl, t_1.\stg, \stg_b)$ and $\I_e(t_1.\stl, t_1.\stg, \stg_a)$. Hence, $t_1.\stg$ provides the existence of $\stg_1$ in the definition of $\R_e$. Thus, $\R_e(t_1.\stl, \stg_a, \stg_b)$. We can use $Hc$ and make the assertion $Q(t^{'}.\stl, \stg_2)$.

@@

Now, we will show that every environment step is in $\R_c$. Consider the step $(t^{'}, \stg_{2k}) \rightarrow_R (t^{'}, \stg_{2(k+1)})$. We have $\I_c(t^{'}.\stl, \stg_2, \stg_{2(k+1)})$ and $\I_c(t^{'}.\stl, \stg_2, \stg_{2k})$. Hence, $\stg_2$ provides the existence of $\stg_1$ in the definition of $\R_c$. Thus, $\R_c(t^{'}.\stl, \stg_{2k}, \stg_{2(k+1)})$.

@@

\begin{theorem}
If 
$$
\forall \stl, \stg, \stg'. \R_e(\stl,\stg,\stg')  \Leftrightarrow  \exists \stg_1.  R(\stg, \stg') \wedge \I_e(\stl, \stg_1, \stg')
$$
and
$$
  \forall \stl,\stg,\stg',\stg''.~
  \neg\I_e(\stl,\stg,\stg') \conj R(\stg',\stg'') \Rightarrow
  \neg\I_e(\stl,\stg,\stg'')
$$
then $\forall \stl, \stg, \stg'.\exists \stg_1. R^{*}(\stg, \stg') \wedge \I_e(\stl, \stg_1, \stg') \Rightarrow \R_e^{*}(\stl, \stg, \stg')$
\end{theorem}
\begin{proof}
Consider $\stl, \stg, \stg', \stg_1$ and $n$ such that $R^{n-1}(\stg, \stg')$ and $\I_e(\stl, \stg_1, \stg')$. Then $R(\stg'_1, \stg'_2)$, $R(\stg'_2, \stg'_3), \ldots, R(\stg'_{n-1}, \stg'_{n})$, where $\stg'_1 = \stg$ and $\stg'_n = \stg'$. We will show that $\R_e^{n-1}(\stl, \stg, \stg')$. Clearly, $\R_e(\stl, \stg'_{n-1}, \stg'_{n})$. Also, $\I_e(\stl, \stg_1, \stg'_{n-1})$. Because otherwise, if $\neg \I_e(\stl, \stg_1, \stg'_{n-1})$, then with $R(\stg'_{n-1}, \stg'_n)$ would imply $\neg \I_e(\stl, \stg_1, \stg'_{n})$ which contradicts our assumption that $\I_e(\stl, \stg_1, \stg')$. Hence, by definition $\R_e(\stl, \stg'_{n-2}, \stg'_{n-1})$. In this manner, we can show that $\forall i, 1 \leq i \leq n-1. \R_e(\stl, \stg'_{i}, \stg'_{i+1})$, because $\I_e(\stl, \stg_1, \stg'_{i+1})$. Hence, $\R_e^{n-1}(\stl, \stg, \stg')$.
\end{proof}

\begin{theorem}
If $\I_e = \I_{ss}$, then $\forall \stl, \stg, \stg'. \R_e(\stl, \stg, \stg') \Rightarrow \stg = \stg'$ 
\end{theorem}
\begin{proof}
First, we show that 
$$
  \forall \stl,\stg,\stg',\stg''.~
  \neg\I_e(\stl,\stg,\stg') \conj R(\stg',\stg'') \Rightarrow
  \neg\I_e(\stl,\stg,\stg'')
$$
Given $\stl, \stg, \stg'$ such that $\neg \I_e(\stl, \stg, \stg')$, $\stg \neq \stg'$. Since $R(\stg', \stg'')$ corresponds to the commit of a transaction, either the transaction is read-only, in which case $\stg' = \stg''$ and hence $\stg \neq \stg''$ which implies $\neg \I_e(\stl, \stg, \stg'')$, or the transaction modifies/inserts a record, in which case it will also add its own unique transaction id to the record, so that $\stg \neq \stg''$, which again implies the result.

Now, consider $\stl, \stg, \stg'$ such that $\R_e(\stl, \stg, \stg')$.By definition of $\R_e$, there exists $\stg_1$ such that $\I_e(\stl, \stg_1, \stg')$. Hence, $\stg_1 = \stg'$. Now, $\I_e(\stl, \stg_1, \stg)$, because otherwise, if $\neg \I_e(\stl, \stg_1, \stg)$, then by the earlier result, $\neg \I_e((\stl, \stg_1, \stg')$ which is a contradiction. Hence, $\stg_1 = \stg$. This implies that $\stg = \stg'$.
\end{proof}

\begin{theorem}
If $\I_e = \I_{ww}$, then $\forall \stl, \stg, \stg'. \R_e(\stl, \stg, \stg') \Rightarrow \stg = \stg'$ 
\end{theorem}

\input{inference-rules}
\newpage
\input{inference-proof}
